---
title: "Portfolio_6"
author: "Lindley Slipetz"
date: "3/15/2021"
output: html_document
---

This project I have to get done quickly because it's time sensitive: it's a March Madness predictor code! The first game is Thursday, so I'll need to get my predictions together before then. We have a lot of datafiles to go through and then I'll try to cobble them together in a way that makes sense to me (and maybe no one else). 

```{r packages_data, warning = FALSE, message = FALSE}
library(tidyverse)
power_rankings <- read.csv("C:\\Users\\Owner\\Documents\\Portfolio_6\\Portfolio_6\\data\\power_rankings.txt", header = TRUE, stringsAsFactors = FALSE)
rpi <- read.csv("C:\\Users\\Owner\\Documents\\Portfolio_6\\Portfolio_6\\data\\rpi.txt", header = TRUE, stringsAsFactors = FALSE)
wab <- read.csv("C:\\Users\\Owner\\Documents\\Portfolio_6\\Portfolio_6\\data\\wab.txt", header = TRUE, stringsAsFactors = FALSE)
```

Okay, now I'm going to make a dataframe that will serve as my bracket. I tried finding something nice online that would 
```{r bracket_creation}
bracket <- data.frame()
team <- c("Gonzaga", "Appalachian St.", "Oklahoma",
           "Missouri", "Creighton", "UCSB",  
           "Virginia",
           "Ohio", "USC", "Wichita St.", "Kansas", 
           "Eastern Wash.", "Oregon", "VCU", "Iowa",
           "Grand Canyon", "Michigan", "Texas Southern", "LSU", "St. Bonaventure", "Colorado",
           "Georgetown", "Florida St.", 
           "UNC Greensboro", "BYU", "UCLA", "Texas",
           "Abilene Christian", "UConn", "Maryland",
           "Alabama", "Iona", "Baylor", "Hartford",
           "North Carolina", "Wisconsin", "Villanova",
           "Winthrop", "Purdue", "North Texas", 
           "Texas Tech", "Utah St.", "Arkansas",
           "Colgate", "Florida", "Virginia Tech",
           "Ohio St.", "Oral Roberts", "Illinois",
           "Drexel", "Loyola Chicago", "Georgia Tech", "Tennessee", "Oregon St.", "Oklahoma St.",
           "Liberty", "San Diego St.", "Syracuse",
           "West Virginia", "Morehead St.",
           "Clemson", "Rutgers", "Houston",
           "Cleveland St.")
bracket <- as.data.frame(cbind(team))
```


