---
title: "Portfolio_6"
author: "Lindley Slipetz"
date: "3/15/2021"
output: html_document
---

This project I have to get done quickly because it's time sensitive: it's a March Madness predictor code! The first game is Thursday, so I'll need to get my predictions together before then. We have a lot of datafiles to go through and then I'll try to cobble them together in a way that makes sense to me (and maybe no one else). 

```{r packages_data, warning = FALSE, message = FALSE}
library(tidyverse)
#install.packages("gdata")
library(gdata)
power_rankings <- read.csv("C:\\Users\\Owner\\Documents\\Portfolio_6\\Portfolio_6\\data\\power_rankings.txt", header = TRUE, stringsAsFactors = FALSE)
rpi <- read.csv("C:\\Users\\Owner\\Documents\\Portfolio_6\\Portfolio_6\\data\\rpi.txt", header = TRUE, stringsAsFactors = FALSE)
basic <- read.csv("C:\\Users\\Owner\\Documents\\Portfolio_6\\Portfolio_6\\data\\basic_stats.txt", header = TRUE, stringsAsFactors = FALSE)
adv <- read.csv("C:\\Users\\Owner\\Documents\\Portfolio_6\\Portfolio_6\\data\\adv_stats.txt", header = TRUE, stringsAsFactors = FALSE)
```

Okay, now I'm going to make a dataframe that will serve as my bracket. I tried finding something nice online that would 
```{r bracket_creation}
bracket <- data.frame(stringsAsFactors = FALSE)
team <- c("Gonzaga", "Appalachian St.", "Oklahoma",
           "Missouri", "Creighton", "UC Santa Barbara",  
           "Virginia",
           "Ohio", "Southern California", "Wichita St.", "Kansas", 
           "Eastern Wash.", "Oregon", "VCU", "Iowa",
           "Grand Canyon", "Michigan", "Texas Southern", "LSU", "St. Bonaventure", "Colorado",
           "Georgetown", "Florida St.", 
           "UNC Greensboro", "BYU", "UCLA", "Texas",
           "Abilene Christian", "UConn", "Maryland",
           "Alabama", "Iona", "Baylor", "Hartford",
           "North Carolina", "Wisconsin", "Villanova",
           "Winthrop", "Purdue", "North Texas", 
           "Texas Tech", "Utah St.", "Arkansas",
           "Colgate", "Florida", "Virginia Tech",
           "Ohio St.", "Oral Roberts", "Illinois",
           "Drexel", "Loyola Chicago", "Georgia Tech", "Tennessee", "Oregon St.", "Oklahoma St.",
           "Liberty", "San Diego St.", "Syracuse",
           "West Virginia", "Morehead St.",
           "Clemson", "Rutgers", "Houston",
           "Cleveland St.")
bracket <- as.data.frame(cbind(team), stringsAsFactors = FALSE)
```

I want to have a dataframe of all the data and a dataframe to use as the bracket. Let's create a new dataframe for all of the data. To do that, I'll need to do some editing of the adv and basic dataframes. The problem is there will be columns with the same name, so we need to rename some columns

```{r ad/ba_redo}
adv$Overall.1[adv$Overall.1 == "W"] <- "W-ovrl"
adv$Overall.2[adv$Overall.2 == "L"] <- "L-ovrl"
adv$Conf.[adv$Conf. == "W"] <- "W-conf"
adv$Conf..1[adv$Conf..1 == "L"] <- "L-conf"
adv$Home[adv$Home == "W"] <- "W-home"
adv$Home.1[adv$Home.1 == "L"] <- "L-home"
adv$Away[adv$Away == "W"] <- "W-away"
adv$Away.1[adv$Away.1 == "L"] <- "L-away"
basic$Overall.1[basic$Overall.1 == "W"] <- "W-ovrl"
basic$Overall.2[basic$Overall.2 == "L"] <- "L-ovrl"
basic$Conf.[basic$Conf. == "W"] <- "W-conf"
basic$Conf..1[basic$Conf..1 == "L"] <- "L-conf"
basic$Home[basic$Home == "W"] <- "W-home"
basic$Home.1[basic$Home.1 == "L"] <- "L-home"
basic$Away[basic$Away == "W"] <- "W-away"
basic$Away.1[basic$Away.1 == "L"] <- "L-away"
colnames(adv) <- as.character(adv[1,])
colnames(basic) <- as.character(basic[1,])
names(adv)[names(adv) == "School"] <- "team"
names(basic)[names(basic) == "School"] <- "team"
adv <- adv[-1, ] 
basic <- basic[-1, ] 
```

Okay, now I'm going to check for repeat columns across the dataframes before combining them. There's also some empty columns in the adv and basic data frames that I'll delete. I'm having trouble deleting them by name, so I'm just going to go with number.


```{r delete_2}
basic <- basic[, -c(1, 3:21)]
adv <- adv[, -c(9, 12, 15, 18, 21)] 
```

Okay, one last thing before we smoosh the dataframes together. We need to make sure the teams match. I've been trying to do this the fancy with agrep, but I've been struggling. Since I'm in a time crunch, I'm just going to do this by brute force. 
```{r name_match}
for(i in 1:nrow(basic)){
    basic[i,1] <- str_replace(basic[i,1], "NCAA", "")
    basic[i,1] <- str_replace(basic[i,1], "State", "St.")
    basic[i,1] <- str_replace(basic[i,1], "UC-Santa Barbara", "UC Santa Barbara")
    basic[i,1] <- str_replace(basic[i,1], "Eastern Washington", "Eastern Wash.")
    basic[i,1] <- str_replace(basic[i,1], "North Carolina-Greensboro", "UNC Greensboro")
}
for(i in 1:nrow(adv)){
    adv[i,1] <- str_replace(adv[i,1], "NCAA", "")
    adv[i,1] <- str_replace(adv[i,1], "State", "St.")
    adv[i,1] <- str_replace(adv[i,1], "UC-Santa Barbara", "UC Santa Barbara")
    adv[i,1] <- str_replace(adv[i,1], "Eastern Washington", "Eastern Wash.")
    adv[i,1] <- str_replace(adv[i,1], "North Carolina-Greensboro", "UNC Greensboro")
}

```
Okay, now let's make a monster dataframe. Okay, after looking at the stats (yes I'm going backwards), I've decided I don't need the basic and adv datasets.

```{r all_data}
NCAA_data <- left_join(bracket, power_rankings, by = "team",  suffix = c(".x", ".y"))
NCAA_data <- left_join(NCAA_data, rpi, by = "team",  suffix = c(".x", ".y"))
```

Okay, now we have our monster dataframe. Before we start doing stuff with it, let's make a quick edit to our bracket dataframe. I want it to represent the different rounds of the tournament, so...

```{r bracket_rev}
brack_names <- c("Round_2", "Round_3",
                       "Round_4", "Round_5", "Round_6", "Round_7")
bracket[ ,brack_names] <- NA

```
Now is the time I step away from Rstudio and start deciding which stats I would like to use to choose my times. From the little I already know about NCAA basketball is that the best offense or (exclusive) the best defense is usually eliminated early. You need a good balance of offense and defense to make it far in the tournament. So two of the measures that I'll be using are a measure of offense and a measure of defense.

First, I'm going to go through all of the columns and see what kind of stat it is.

# Stat list

* __yusag coeefficient__:  number of points they're expected to win by over average team 

* ___rpi__: measure of wins, losses, strength of schedule __will use rpi_rank for this__

* __offense coefficient__: the creator of this dataset is not clear on exactly what went into creating this coefficient, but the higher the number, the better the offense 

* __defense coefficient__: the creator of this dataset is not clear on exactly what went into creating this coefficient, but the higher the number, the better the defense 

The yusag coefficient seems to be a combination of defense and offense coefficients so I think it would be redundant to use it. I want to pit the offense and the defense stats together, but I think it's important to consider strength of schedule with that. I'm going to make two new columns: off_sos_rank and def_sos_rank, which will be the rankings of offense and defense with strength of schedule considered.

```{r off_sos_rank}
NCAA_data <- NCAA_data %>%
  mutate(
    off_sos_rank =  as.numeric(rank((off_coeff*rpi)))
  )
```

```{r def_sos_rank}
NCAA_data <- NCAA_data %>%
  mutate(
    def_sos_rank =  as.numeric(rank((def_coeff*rpi)))
  )
```

Okay, cool. Now we have all the data we need. Now I'm going to write a function to compare teams. Okay, nevermind. My function is not working, so I'm just going to throw it all in a while loop.

```{r compare_func}
compare_teams <- function(dataframe_data, dataframe_bracket, round, team1, team2, offense, defense) {

}
```

Cool, I have my function. Now, let's put it to use.

```{r round_2}
count <- 1
round <- 2
while(count < 65){
  second <- count + 1
  team1 <- bracket$team[count]
  team2 <- bracket$team[second]
  offense_1 <- NCAA_data$off_sos_rank[count]
  offense_2 <- NCAA_data$off_sos_rank[second]
  defense_1 <- NCAA_data$def_sos_rank[count]
  defense_2 <- NCAA_data$def_sos_rank[second]
  compare_1 <- offense_1 - defense_2
  compare_2 <- defense_1 - offense_2
  result = compare_1 + compare_2
  if(result > 0){
    bracket <- bracket %>%  
      add_row(Round_2 = team1)
      }
   else{
     bracket <- bracket %>%  
      add_row(Round_2 = team2)
      }
  count <- count + 2
}
```

Okay, that got me undesirable results. Let's try again with different stats.

```{r round_2b}
count <- 1
round <- 2
while(count < 65){
  second <- count + 1
  team1 <- bracket$team[count]
  team2 <- bracket$team[second]
  offense_1 <- NCAA_data$off_rank[count]
  offense_2 <- NCAA_data$off_rank[second]
  defense_1 <- NCAA_data$def_rank[count]
  defense_2 <- NCAA_data$def_rank[second]
  compare_1 <- offense_1 - defense_2
  compare_2 <- defense_1 - offense_2
  result = compare_1 + compare_2
  if(result > 0){
    bracket <- bracket %>%  
      add_row(Round_2 = team1)
      }
   else{
     bracket <- bracket %>%  
      add_row(Round_2 = team2)
      }
  count <- count + 2
}
```

That goes in the opposite direction (the problem with the first one was that only the higher-seeded team won; the problem with the second one is that now the lower-seeded teams are winning). Let's try a new statistic. Like I previously mentioned, the best teams are a balance of good offense and defense. Let's take the average of offense and defense rank to find the balance.

```{r all_rank}
NCAA_data <- NCAA_data %>%
  mutate(
    all_rank =  as.numeric(rank((off_rank+def_rank)/2))
  )
```

```{r round_2c}
count <- 1
while(count < 65){
  second <- count + 1
  team1 <- bracket$team[count]
  team2 <- bracket$team[second]
  all_1 <- NCAA_data$all_rank[count]
  all_2 <- NCAA_data$all_rank[second]
  result = all_1 - all_2
  if(result > 0){
    bracket <- bracket %>%  
      add_row(Round_2 = team1)
      }
   else{
     bracket <- bracket %>%  
      add_row(Round_2 = team2)
      }
  count <- count + 2
}
```

Okay, I think 2b and 2c give the same results (and maybe that's obvious...I'm moving fast through this, so I'm not really thinking of the implications). Let's run it for the other rounds. Since I'm in two pools, I think I'll do two different brackets: one for 1a and one for 1c.

```{r round_3c}
count <- 1
while(count < 33){
  second <- count + 1
  team1 <- bracket$Round_2[64 + count]
  team2 <- bracket$Round_2[64 + second]
  all_1 <- NCAA_data$all_rank[count]
  all_2 <- NCAA_data$all_rank[second]
  result = all_1 - all_2
  if(result > 0){
    bracket <- bracket %>%  
      add_row(Round_3 = team1)
      }
   else{
     bracket <- bracket %>%  
      add_row(Round_3 = team2)
      }
  count <- count + 2
}
```

```{r round_4c}
count <- 1
while(count < 17){
  second <- count + 1
  team1 <- bracket$Round_3[96 + count]
  team2 <- bracket$Round_3[96 + second]
  all_1 <- NCAA_data$all_rank[count]
  all_2 <- NCAA_data$all_rank[second]
  result = all_1 - all_2
  if(result > 0){
    bracket <- bracket %>%  
      add_row(Round_4 = team1)
      }
   else{
     bracket <- bracket %>%  
      add_row(Round_4 = team2)
      }
  count <- count + 2
}
```

```{r round_5c}
count <- 1
while(count < 9){
  second <- count + 1
  team1 <- bracket$Round_4[112 + count]
  team2 <- bracket$Round_4[112 + second]
  all_1 <- NCAA_data$all_rank[count]
  all_2 <- NCAA_data$all_rank[second]
  result = all_1 - all_2
  if(result > 0){
    bracket <- bracket %>%  
      add_row(Round_5 = team1)
      }
   else{
     bracket <- bracket %>%  
      add_row(Round_5 = team2)
      }
  count <- count + 2
}
```

Okay, this is not working. If you know anything about basketball, and you see the final four I have, you'll see why this is NOT a good predictor. I'm going back to 1a.

```{r round_3a}
count <- 1
round <- 2
while(count < 33){
  second <- count + 1
  team1 <- bracket$Round_2[64 + count]
  team2 <- bracket$Round_2[64 + second]
  offense_1 <- NCAA_data$off_sos_rank[count]
  offense_2 <- NCAA_data$off_sos_rank[second]
  defense_1 <- NCAA_data$def_sos_rank[count]
  defense_2 <- NCAA_data$def_sos_rank[second]
  compare_1 <- offense_1 - defense_2
  compare_2 <- defense_1 - offense_2
  result = compare_1 + compare_2
  if(result > 0){
    bracket <- bracket %>%  
      add_row(Round_3 = team1)
      }
   else{
     bracket <- bracket %>%  
      add_row(Round_3 = team2)
      }
  count <- count + 2
}
```

```{r round_4a}
count <- 1
round <- 2
while(count < 17){
  second <- count + 1
  team1 <- bracket$Round_3[96 + count]
  team2 <- bracket$Round_3[96 + second]
  offense_1 <- NCAA_data$off_sos_rank[count]
  offense_2 <- NCAA_data$off_sos_rank[second]
  defense_1 <- NCAA_data$def_sos_rank[count]
  defense_2 <- NCAA_data$def_sos_rank[second]
  compare_1 <- offense_1 - defense_2
  compare_2 <- defense_1 - offense_2
  result = compare_1 + compare_2
  if(result > 0){
    bracket <- bracket %>%  
      add_row(Round_4 = team1)
      }
   else{
     bracket <- bracket %>%  
      add_row(Round_4 = team2)
      }
  count <- count + 2
}
```


```{r round_5a}
count <- 1
round <- 2
while(count < 9){
  second <- count + 1
  team1 <- bracket$Round_4[112 + count]
  team2 <- bracket$Round_4[112 + second]
  offense_1 <- NCAA_data$off_sos_rank[count]
  offense_2 <- NCAA_data$off_sos_rank[second]
  defense_1 <- NCAA_data$def_sos_rank[count]
  defense_2 <- NCAA_data$def_sos_rank[second]
  compare_1 <- offense_1 - defense_2
  compare_2 <- defense_1 - offense_2
  result = compare_1 + compare_2
  if(result > 0){
    bracket <- bracket %>%  
      add_row(Round_5 = team1)
      }
   else{
     bracket <- bracket %>%  
      add_row(Round_5 = team2)
      }
  count <- count + 2
}
```

```{r round_6a}
count <- 1
round <- 2
while(count < 5){
  second <- count + 1
  team1 <- bracket$Round_5[120 + count]
  team2 <- bracket$Round_5[120 + second]
  offense_1 <- NCAA_data$off_sos_rank[count]
  offense_2 <- NCAA_data$off_sos_rank[second]
  defense_1 <- NCAA_data$def_sos_rank[count]
  defense_2 <- NCAA_data$def_sos_rank[second]
  compare_1 <- offense_1 - defense_2
  compare_2 <- defense_1 - offense_2
  result = compare_1 + compare_2
  if(result > 0){
    bracket <- bracket %>%  
      add_row(Round_6 = team1)
      }
   else{
     bracket <- bracket %>%  
      add_row(Round_6 = team2)
      }
  count <- count + 2
}
```

```{r round_7a}
count <- 1
round <- 2
while(count < 3){
  second <- count + 1
  team1 <- bracket$Round_6[124 + count]
  team2 <- bracket$Round_6[124 + second]
  offense_1 <- NCAA_data$off_sos_rank[count]
  offense_2 <- NCAA_data$off_sos_rank[second]
  defense_1 <- NCAA_data$def_sos_rank[count]
  defense_2 <- NCAA_data$def_sos_rank[second]
  compare_1 <- offense_1 - defense_2
  compare_2 <- defense_1 - offense_2
  result = compare_1 + compare_2
  if(result > 0){
    bracket <- bracket %>%  
      add_row(Round_7 = team1)
      }
   else{
     bracket <- bracket %>%  
      add_row(Round_7 = team2)
      }
  count <- count + 2
}
```